---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

Read data file using read.csv() function, then convert to data table.

```{r warning=FALSE}
library(data.table)
library(ggplot2)
library(plyr)
library(dplyr)

dt_act<-data.table(read.csv("activity.csv"))


```

## What is mean total number of steps taken per day?


First we calculate the total number of steps per day, then get the summary of the totoal steps.
```{r }

dt_totalsteps<-dt_act[,sum(steps,na.rm = TRUE),by=date]

names(dt_totalsteps)<-c("date","total_steps")

data_summary<-summary(dt_totalsteps$total_steps)
print(data_summary)
```
The mean value of total steps per day is `r data_summary[4]` and the median value of total  steps per day is `r format(data_summary[3],scientific=FALSE)`


Here is the histogram of total steps per day with bins=40:

```{r  warning=FALSE}

hist_totalsteps<-ggplot(data=dt_totalsteps,aes(total_steps))
hist_totalsteps<-hist_totalsteps+geom_histogram(bins = 40)+xlab("total steps")
hist_totalsteps<-hist_totalsteps+ggtitle("Total steps per day")
print(hist_totalsteps)

```

## What is the average daily activity pattern?

We first calculate the average number of steps vs. interval, then plot the time series plot of the 5-minute interval and the average number of steps taken across all days.

```{r  warning=FALSE}


dt_act[,step_mean:=mean(steps,na.rm=TRUE),by=interval]

g<-ggplot(data=dt_act,aes(interval,step_mean))
g<-g+geom_line()+ylab("Average steps")+ggtitle("Average steps vs. interval")
print(g)
```

Then we calculate the maximum number of steps averaged across all the days:
```{r}

print(c(max(dt_act$step_mean),dt_act$interval[which.max(dt_act$step_mean)]))

```

The maximum number is `r max(dt_act$step_mean)` and the 5 minute inverval with this maximum number is `r dt_act$interval[which.max(dt_act$step_mean)]`

## Imputing missing values

The total number of missing value in the data set is `r sum(is.na(dt_act$steps))` . 

Here we use the mean value of steps to replace the "NA" missing value in steps, and saving the original steps value into "old_steps".

A new dataset "dt_new_act" is created. It is equal to the original dataset but with the missing data filled in.

```{r warning=FALSE, results="hide" }
dt_act[,old_steps:=steps]

dt_act[,steps:=replace(steps,is.na(steps),step_mean)]

dt_new_act<-select(dt_act,steps,date,interval)


```

We calculate the total steps in the the new data set "dt_new_act" per day and plot the histogram 

```{r}
dt_new_totalsteps<-dt_new_act[,sum(steps),by=date]
names(dt_new_totalsteps)<-c("date","new_total_steps")


data_summary_2<-summary(dt_new_totalsteps$new_total_steps)
print(data_summary_2)

hist_new_totalsteps<-ggplot(data=dt_new_totalsteps,aes(new_total_steps))
hist_new_totalsteps<-hist_new_totalsteps+geom_histogram(bins=40)+xlab("total steps")+ggtitle("Total steps per day with filled data")
print(hist_new_totalsteps)

```

The mean value of total steps per day is `r format(data_summary_2[4],scientific=FALSE)` and the median value of total  steps per day is `r format(data_summary_2[3],scientific=FALSE)`. the difference between those values comparing with first part is `r data_summary_2[4]-data_summary[4]` and `r data_summary_2[3]-data_summary[3]`. The filled data increased bothe the mean value and the median value.



## Are there differences in activity patterns between weekdays and weekends?

First we split teh data from weekdays and weekends.

```{r}


allweekdays= c("Monday","Tuesday","Wednesday","Thursday","Friday")

dt_new_act$weekday<-as.factor(ifelse(weekdays(as.Date(dt_new_act$date))%in%allweekdays,"weekday","weekend"))

weekdaysplit<-ddply(dt_new_act, c("interval","weekday"),summarise,
                                     meansteps = mean(steps,na.rm=TRUE)
)

weekdayplot<-ggplot(weekdaysplit,aes(x=interval,y=meansteps))+
       facet_wrap(~weekday,nrow=2,ncol=1)+ geom_line()
 weekdayplot<-weekdayplot+ ggtitle("Mean steps over each 5min interval split by weekday/weekend")+ylab("Average steps")+ xlab("Interval ")
print(weekdayplot)


```

So in weekends, the activity is higher in weekedn that in weekday after 1000 intervals 